---
title: "Percentage of Parents'Categories"
author: "Keshav Bhusal"
date: "2022-10-11"
output:
html_document: default
---

### Setting up the environment

Install the `tidyverse` package `install.packages("tidyverse")`

```{r loading packages }
   library(tidyverse)
   library(dplyr)
   library(readr)
   
```

### Importing the data set

Load the BRFSS_2020 data file [BRFSS_2020.csv](https://arizona.app.box.com/file/1015177716447?s=4s9owh47wyny08u9tkui6h34pmek4kig)

### Creating the dataframe with required variables

```{r}
limited_dataframe <- BRFSS_2020%>%
  select("_SEX", BIRTHSEX, CHILDREN, MARITAL) %>% 
  rename(SEX = "_SEX")
```

### Transformation of variables into groups

```{r}
limited_dataframe$MARITAL[limited_dataframe$MARITAL==2]<-"single_parent"
limited_dataframe$MARITAL[limited_dataframe$MARITAL==3]<-"single_parent"
limited_dataframe$MARITAL[limited_dataframe$MARITAL==4]<-"single_parent"
limited_dataframe$MARITAL[limited_dataframe$MARITAL==5]<-"single_parent"
limited_dataframe$MARITAL[limited_dataframe$MARITAL==1]<-"double_parent"
limited_dataframe$MARITAL[limited_dataframe$MARITAL==6]<-"double_parent"
limited_dataframe$MARITAL[limited_dataframe$MARITAL==9]<-"refused"
limited_dataframe$CHILDREN[limited_dataframe$CHILDREN==88]<-0
limited_dataframe$CHILDREN[limited_dataframe$CHILDREN==99]<-NA
limited_dataframe$SEX[limited_dataframe$SEX==2]<-"female"
limited_dataframe$SEX[limited_dataframe$SEX==1]<-"male"
```

### Creating a new variable with parents categorization

```{r}
new_column <- limited_dataframe %>% 
  mutate(parent_type = case_when(MARITAL == "single_parent" & SEX == "male" ~ "father_only",
                             MARITAL == "single_parent" & SEX == "female"  ~ "mother_only", 
                             MARITAL == "refused" ~ "refused", 
                             MARITAL == "double_parent" ~ "double_parent"))
```

### Creating the smaller required dateframe

```{r}
dataframe <- new_column %>%
  select(CHILDREN, parent_type) 
```

### Grouping and aggregating the categories

```{r}
categorical <- dataframe %>%
  group_by(parent_type) %>%
  summarise(CHILDREN = sum(CHILDREN, na.rm = TRUE)) %>% 
  drop_na(parent_type)
```

### Calculating the percentage of parents category

```{r}
  percentage <- categorical%>%
  mutate(total= sum(CHILDREN)) %>% 
  mutate(percentage=CHILDREN/total*100)
```
