---
title: "Merging the BRFSS data files"
author: "Keshav Bhusal"
date: "2022-10-18"
output:html_document: default
---

### Setting up the environment

Install the `tidyverse` package `install.packages("tidyverse")` Install the `rmarkdown` package `install.packages("rmarkdown")`

```{r loading packages }
   library(tidyverse)
   library(dplyr)
   library(readr)
   library(rmarkdown)
   
```

### Importing the data set

Loading all the BRFSS data file from 2017 to 2021

[brfss2017](D:/RA_Works/DATA_SETS/BRFSS_2017.csv)
[brfss2018](D:/RA_Works/DATA_SETS/BRFSS_2018.csv)
[brfss2019](D:/RA_Works/DATA_SETS/BRFSS_2019.csv)
[brfss2020](D:/RA_Works/DATA_SETS/BRFSS_2020.csv)
[brfss2021](D:/RA_Works/DATA_SETS/BRFSS_2021.csv)

```{r}
brfss2017 <- read_csv("D:/RA_Works/DATA_SETS/BRFSS_2017.csv")
brfss2018 <- read_csv("D:/RA_Works/DATA_SETS/BRFSS_2018.csv")
brfss2019 <- read_csv("D:/RA_Works/DATA_SETS/BRFSS_2019.csv")
brfss2020 <- read_csv("D:/RA_Works/DATA_SETS/BRFSS_2020.csv")
brfss2021 <- read_csv("D:/RA_Works/DATA_SETS/BRFSS_2021.csv")

```

### Merging the multi-year data files

First observing the **common columns** occurring in both years **2017** and **2018**

```{r}
brfss1718 <- intersect(colnames(brfss2017), colnames(brfss2018))
brfss1718
```

Now, combining/binding the data files of the years **2017** and **2018** on the given common columns.

```{r}
brfss17_18 <- rbind(brfss2017[ , brfss1718], brfss2018[ , brfss1718])
```

```{r}
glimpse(brfss17_18)
```

Similarly, intersecting and combining/binding the data files of the years **2019** and **2020**.

```{r}
brfss1920 <- intersect(colnames(brfss2019), colnames(brfss2020))
brfss1920
```

```{r}
brfss19_20 <- rbind(brfss2019[ , brfss1920], brfss2020[ , brfss1920])
```

```{r}
glimpse(brfss19_20)
```

Similarly, intersecting and combining/binding the data files of the years **2017-2018** and **2019-2020**.

```{r}
brfss1720 <- intersect(colnames(brfss17_18), colnames(brfss19_20))
brfss1720
```

```{r}
brfss17_20 <- rbind(brfss17_18[ , brfss1720], brfss19_20[ , brfss1720])
```

```{r}
glimpse(brfss17_20)
```

Finally, intersecting and combining/binding the data files of the years **2017-2020** and **2021**.

```{r}
brfss_all <- intersect(colnames(brfss17_20), colnames(brfss2021))
brfss_all
```

```{r}
brfss_all_17_21 <- rbind(brfss17_20[ , brfss_all], brfss2021[ , brfss_all])
```

```{r}
glimpse(brfss_all_17_21)
```

Thus, 'brfss_all_17_21' is the final merged data file with **147** variables and **2146371** observations.

### Creating the dataframe with required variables

I have hereby created the data frame of the required variables for the calculation of **percentage of single parents**. I took the `CHILDREN` and `MARITAL` variables.

I could not account for the gender of the single parent because `_SEX` (a calculated variable), is was not commonly available in all the year's data sets and so it got eventually discarded.

```{r}
limited_dataframe <- brfss_all_17_21 %>%
  select(CHILDREN, MARITAL)
```

### Transformation of variables into groups

Based on the area of interest and making things look clear and comfortable, I just transformed the responses to the categories.

```{r}
limited_dataframe$MARITAL[limited_dataframe$MARITAL==2]<-"single_parent"
limited_dataframe$MARITAL[limited_dataframe$MARITAL==3]<-"single_parent"
limited_dataframe$MARITAL[limited_dataframe$MARITAL==4]<-"single_parent"
limited_dataframe$MARITAL[limited_dataframe$MARITAL==5]<-"single_parent"
limited_dataframe$MARITAL[limited_dataframe$MARITAL==1]<-"double_parent"
limited_dataframe$MARITAL[limited_dataframe$MARITAL==6]<-"double_parent"
limited_dataframe$MARITAL[limited_dataframe$MARITAL==9]<-"refused"
limited_dataframe$CHILDREN[limited_dataframe$CHILDREN==88]<-0
limited_dataframe$CHILDREN[limited_dataframe$CHILDREN==99]<-NA
```

### Grouping and aggregating the categories

```{r}
categorical <- limited_dataframe %>%
  group_by(MARITAL) %>%
  summarise(CHILDREN = sum(CHILDREN, na.rm = TRUE)) %>% 
  drop_na(MARITAL)
```

### Calculating the percentage of parents category

```{r}
  percentage <- categorical%>%
  mutate(total= sum(CHILDREN)) %>% 
  mutate(percentage=CHILDREN/total*100)
```
